### 2016-05-23 paper.tex

\documentclass[10pt,twocolumn]{article}
\usepackage{times}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx,grffile}
\usepackage{pgf}
\usepackage{tikz}
\usepackage{listings}
\usepackage{color}
\usetikzlibrary{arrows,automata}
%\usepackage[latin1]{inputenc}

\lstset{frame=tb,
  language=Ruby,
  captionpos=b,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\scriptsize\ttfamily},
  numbers=none,
  %numberstyle=\tiny\color{gray},
  %keywordstyle=\color{blue},
  %commentstyle=\color{dkgreen},
  %stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

% do not change these values
\baselineskip 12pt
\textheight 9in
\textwidth 6.5in
\oddsidemargin 0in
\topmargin 0in
\headheight 0in
\headsep 0in

%\makeatletter
%\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
%\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
%\makeatother
%% Scale images if necessary, so that they will not overflow the page
%% margins by default, and it is still possible to overwrite the defaults
%% using explicit options in \includegraphics[width, height, ...]{}
%\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
%

\usepackage[unicode=true]{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\usepackage[font=footnotesize,labelfont=bf]{caption}

\begin{document}

\title{outline!}

\author{
%Noah Watkins, Neha Ojha\textsuperscript{*}, Carlos Maltzahn \\
%\small {\em University of California, Santa Cruz} \\
%\small {\{jayhawk,carlosm\}@cs.ucsc.edu} \textsuperscript{*}nojha@ucsc.edu \\ [2mm]
\small Submission Type: Research
}

\date{}
\maketitle

\begin{abstract}
To meet the needs of a diverse and growing set of cloud-based applications,
modern distributed storage frameworks expose a variety of composable
subsystems as building blocks.  This approach gives infrastructure programmers
significant flexibility in implementing application-specific semantics while
reusing trusted components.  Unfortunately, in current storage systems the
composition of subsystems is a low-level task that couples (and hence
obscures) a variety of orthogonal concerns, including functional correctness
and performance.  Building an application by wiring together a collection of
components typically requires thousands of lines of carefully-written C++
code, an effort that must be repeated whenever device or subsystem
characteristics change.

In this paper, we propose a declarative approach to subservice composition
that allows programmers to focus on the high-level functional properties that
are required by applications.  Choosing an implementation that is consistent
with the declarative functional specification then can be posed as a search
problem over the space of parameters such as block sizes, storage interfaces
(e.g. key/value or block storage) and concurrency control mechanisms.  We
present experimental evaluation of our prototype, (etc etc)
\end{abstract}

\section{Introduction}

% whats the problem
Storage systems are increasingly providing features that take advantage of
application-specific knowledge to achieve optimizations and provide unique
services. However, this trend is leading to the creation of a large number of
software extensions that will be difficult to maintain as system software and
hardware continue to evolve.

%%% Anderson FAST 2016 IoT Time Series DB on Ceph

% why is it interesting
The standardization of the POSIX file I/O interface has been a major success,
allowing application developers to avoid vendor lock-in. However, large-scale
storage systems have been dominated by proprietary products, preventing
exploration of alternative interfaces and complicating future migration paths,
eliminating the benefits of commodity systems. But the recent availability of
high-performance open-source storage systems is changing this because these
systems are modifiable, enabling interface change, and reducing the risks of
lock-in. The widely deployed Ceph distributed storage system is an example of
a storage system that supports application-specific extensions in the form of
custom I/O interfaces to objects managed by the underlying RADOS object
storage system~\cite{weil:osdi06,weil:pdsw07}. Organizations are increasingly
reliant upon these extensions as is shown in Figure~\ref{fig:objclass-dev} by
a marked increase in the number of object operations that are packaged as part
of the Ceph distribution and widely used by internal Ceph subsystems and by
applications such as OpenStack Swift and Cinder~\cite{openstack}.

\begin{figure}[t]
  \centering
    \includegraphics[width=0.48\textwidth]{experiments/objclass-dev/output.png}
    \caption{
[\href{https://github.com/noahdesu/zlog-popper/tree/master/experiments/objclass-dev/visualize.ipynb}{source}]
Growth of officially supported, custom object interfaces in RADOS over 6
years. An \emph{operation} is a function executed in the context of an object,
and operations are grouped into different \emph{categories}
corresponding to applications or utilities, such as reference counting}
\label{fig:objclass-dev}
\end{figure}

%%% why is it hard / why do naive approaches fail
In addition to the growth in the quantity of operations in use throughout Ceph
installations, Figure~\ref{fig:objclass-dev} also depicts the amount of
low-level C++ written to implement these operations. Unfortunately, this code
is written assuming a performance profile defined by the combination of the
hardware and software versions available at the time of development.  While
the bulk of these interfaces are created by core Ceph developers with a
complete view of the performance model, this may be changing as the
development community has been receptive to outside contributions with the
recent inclusion by CERN developers of an extension for performing limited
numeric operations on object data~\cite{cls_numops}. And while Ceph has not
yet reached the point of directly exposing these features to
non-administrative users, the recent inclusion of a mechanism for dynamically
defining extensions using Lua~\cite{cls_lua} suggests that aspects of this
feature may soon appear. What is needed is support for creating storage
interfaces using a method that allows transparent optimization as the system,
application, and supporting environment evolve.

% why hasn't it been solved before / why have previous approaches failed
Previous work related to storage interface design has largely been in the
context of standardization efforts and active storage. While the later has
been specifically concerned with the creation of application-specific
interfaces and demonstrating short-term benefits, the standardization seeks to
develop a last set of interfaces. In neither case have we seen efforts focused
on extensible interfaces with portability and longevity a primary concern.

%This trend is also not limited to Ceph. AWS Lambda, Redis Lua, Redis Modules,
%KV-Drives, and Rhea, are all recent examples of storage systems embracing the
%power of including application-specific semantics.

We propose the introduction of a declarative language for building object
based interfaces that allows a storage system to meet the needs of an
application throughout the development process without requiring rewrites.

\section{Background and Motivation}

In this section we highlight the salient components of Ceph, especially its
\emph{object class} feature that offers users the ability to load and execute
application-specific codes. We provide a description of our motivating
example, a high-performance distributed shared-log built upon Ceph that makes
extensive use of the object class facility.

\subsection{Ceph and Storage Programmability}
\label{sec:objclass}

Figure~\ref{fig:ceph} illustrates the collection of components commonly
referred to as Ceph. At the bottom, a cluster of 10s--10,000s \emph{object
storage devices} compose the distributed object storage system called RADOS.
Widely deployed applications such as the S3/Swift--compliant RADOS Gateway
(RGW), RADOS Block Device (RBD), and the POSIX Ceph File System are built upon
the \emph{librados} client layer that presents a fault-tolerant always-on view
of the RADOS cluster.

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{.48\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{figures/ceph}
      \caption{Ceph}
      \label{fig:ceph}
  \end{subfigure}\quad
  \begin{subfigure}[b]{.40\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{figures/osd}
      \caption{OSD}
      \label{fig:osd}
  \end{subfigure}
  \caption{Ceph and OSD... gotta cram more stuff in here...}
\end{figure}

The object storage device (OSD), illustrated in Figure~\ref{fig:osd}, is the
building block of the RADOS cluster and is responsible for managing and
providing access to a set of
named objects. The configuration of an OSD is flexible, and commonly contains
a mix of commodity hardware such as HDD and SSD bulk storage, a multi-core
CPU, GBs of RAM, and one or two 10 Gb Ethernet links. Clients access object
data managed by an OSD by invoking native object operations exposed by the OSD
such as reading or writing bytes, as well as more complex operations like taking
snapshots or composing one or more native operations into compound procedures
that execute in a transactional context.

The native object operations in RADOS roughly fall into two categories based
on the type of data being accessed: key-value items, or bulk bytestream data.
The key-value interface operates as a dedicated database associated with each
object, and the bytestream interface supports random byte-level access similar
to a file.  At a
low-level each of these abstract I/O interfaces map to hardware storage
devices through a pluggable object backend storage service. For instance,
LevelDB or RocksDB may be used to store key-value data, while the
\emph{FileStore} implementation maps the bytestream interface onto a local
POSIX file system~\cite{leveldb,rocksdb}. Several backend implementations
exist for storing data in targets such as the Kinetic Drive or in NVMe devices,
among others.

%Independent of the hardware or backends chosen, Ceph allows all
%object operations to be combined to create compound operations such as storing
%key-value metadata associated with a binary blob, allowing applications to
%build object interfaces that work even as Ceph software and hardware evolve.

\paragraph*{Object Classes}
While Ceph provides a wide variety of native object operations, it also
includes a facility referred to as \emph{object classes} that allow developers
to create application-specific object operations in the form of C++ shared
libraries dynamically loaded into the OSD process at runtime.  Object classes
can be used to implement basic data management tasks such as indexing
metadata, or used to perform complex operations such as data transformations
or filtering. Table~\ref{tab:objclass-cats} summarizes the range of object
classes maintained in the upstream Ceph project which support internal Ceph
subsystems as well as applications and services that run on top of Ceph.

\begin{table}[ht]
\centering
\begin{tabularx}{\columnwidth}{|X|l|l|l|}
\hline
Category & Specialization & Methods \\ \hline
\multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
                         & Exclusive & \\ \hline
\multirow{3}{*}{Logging} & Replica & 3 \\
                         & State & 4 \\
                         & Timestamped & 4 \\ \hline
Garbage Collection & Ref. Counting & 4 \\ \hline
\multirow{4}{*}{Metadata} & RBD & 37 \\
 & RGW & 27 \\
 & User & 5 \\
 & Version & 5 \\ \hline
\end{tabularx}
\caption{A variety of RADOS object storage classes exist that expose reusable
interfaces to applications.}
\label{tab:objclass-cats}
\end{table}

A critical step in the development of application-specific object interfaces
is deciding how to best make use of the native object interfaces. For instance
if an application stores an image in an object, it may also extract and store
EXIF metadata as key-value pairs in the object key-value database.  However,
depending on the application needs it may be sufficient or offer a performance
advantage to store this metadata as a header within the bytestream. In the remainder
of this section we will explore the challenges associated with these design questions.

\subsection{Motivating Application: CORFU}

The primary motivating example we will use in this paper is the CORFU
distributed shared-log designed to provide high-performance serialization
across a set of flash storage devices~\cite{balakrishnan:nsdi12}. The
shared-log is a powerful abstraction useful when building distributed systems
and applications, but common implementations such as Paxos or Raft funnel I/O
through a single node limiting total throughput~\cite{lamport:tocs89}. The CORFU protocol addresses this limitation
by de-coupling log entry storage from log metadata management, making use of a
centralized, volatile, in-memory \emph{sequencer service} that assigns
positions to clients that are appending to the log. Since the sequencer is
centralized serialization is trivial, and the use of non-durable state allows
the sequencer service to operate at very high rates. The CORFU system has been
used to demonstrate a number of interesting services such as transactional
key-value and metadata services, replicated state machines, and an elastic
cloud-based database management system~\cite{balakrishnan:sosp13,bernstein:cidr11}.

Two aspects of CORFU make its design attractive in the context of the Ceph
storage system. First, CORFU assumes a cluster of flash devices because
log-centric systems tend to have a larger percentage of random reads making it
difficult to achieve high-performance with spinning disks. However, the speed
of the underlying storage does not affect correctness. Thus, in a
software-defined storage system such as Ceph a single implementation can
transparently take advantage of any software or hardware upgrades, and make use
of existing and future data management features such as tiering,
allowing users to freely choose between media types such as SSD, spinning
disks for archival storage, or emerging NVRAM technologies.

\paragraph*{CORFU and Storage Programmability}
The second property of CORFU relevant in the context of Ceph is the dependency
CORFU places on custom storage device interfaces used to guarantee
serialization during failure and reconfiguration. Each flash device in a CORFU
cluster exposes a 64-bit write-once address space consisting of the primary I/O
interfaces \emph{write(pos, data)} and \emph{read(pos)} for accessing log
entries, as well as \emph{fill(pos)} and \emph{trim(pos)} that invalidate and
reclaim log entries, respectively. All I/O operations in CORFU initiated by
clients are tagged with an \emph{epoch} value, and flash devices are expected
to reject client requests that contain an old epoch value. To facilitate
recovery or handle system reconfiguration in CORFU, the storage devices are
also required to support a \emph{seal(epoch)} command that stores the latest
epoch and returns the maximum position written to that device. The seal
interface is used following the failure of a sequencer to calculate the tail of
the log that the sequencer should use to repopulate its in-memory state.

While the authors of the CORFU paper describe prototype device interfaces
implemented as both host-based and FPGA-based solutions, RADOS \emph{directly} supports
the creation of logical storage devices through its object class feature
described previously in Section~\ref{sec:objclass}. Thus, by using
software-based object interfaces offered by RADOS flash devices in CORFU can be
replaced by software-defined storage offering significant flexibility and a
simplified design.

The implementation of a custom object class that satisfies the needs of an
application such as CORFU is often straightforward. However, as described in
Section~\ref{sec:objclass} there are a variety of native object I/O interfaces
available, and it is not always immediately clear how best to utilize these
interfaces.

\section{Physical Design}

As we have seen, Ceph provides a rich storage API and places few restrictions
on the structure of applications. Thus a primary concern when implementing an
object interface in Ceph is deciding how native interfaces are composed into
compound operations in order to implement the semantics of the target
interface. These types of decisions are commonly referred to as
physical design, and can affect performance and application flexibility.
As we will see the design space that developers must operate in is often
large, and its dynamic nature can lead to design decisions that become
obsolete and non-optimal.

\begin{figure}[t]
\centering
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.2cm,semithick]
%\tikzstyle{every state}=[fill=red,draw=none,text=white]

  \node[initial left,state] (A)              {$EG$};
  \node[state]         (B) [below right of=A] {$CP$};
  \node[state]         (D) [right of=B]       {$IO$};
  \node[state]         (C) [above right of=A] {$GC$};
  \node[state]         (E) [right of=C]       {$US$};
  \node[state]         (F) [below right of=E] {$Fi$};

  \path (A) edge        node [left] {R,W,F} (B)
            edge        node {T}     (C)
        (B) edge        node {F}     (C)
            edge        node {R,W}   (D)
        (C) edge        node {F,T}   (E)
        (D) edge        node {W}     (E)
            edge        node {R}     (F)
        (E) edge        node {F,T,W} (F);
\end{tikzpicture}
\caption{State transition diagram for read ($R$), write ($W$), fill ($F$), and
trim ($T$) CORFU operations. The states epoch guard ($EG$), check position ($CP$),
and update state ($US$) access metadata. The I/O performs a log entry read or
write, and garbage collection ($GC$) marks entries for reclamation.}
\label{fig:corfu-sm}
\end{figure}

To understand the developer process in the context of CORFU we have included the
state-machine diagram in Figure~\ref{fig:corfu-sm} showing the composition of
actions for each component of the CORFU interface which must be mapped onto
Ceph object classes. For instance, all operations begin by applying an
\emph{epoch guard} that ensures the request is tagged with an up-to-date epoch
value. The \emph{read} (R) and \emph{write} (W) operations both proceed by (1)
examining metadata associated with the target position, (2) performing I/O to
read or write the log entry, and in the case of a write, (3) updates metadata
for the target log position.

As an example, one valid design option is to store each log position in an
object with a name using a one-to-one mapping with the log entry position.
This would simplify the design of the \emph{write} interface because a small
amount of metadata stored as a header in the object could describe the state
of the log entry.  However, as we will see this choice of a physical design
can result in poor performance compared to other designs. In the remainder of
this section we will define the entire design space and use a set of targetted
benchmarks to arrive at a final design. Finally we will show that these design
decisions can lead to non-optimal decisions and suggest an automated approach
is desirable.

\subsection{Challenges}

The design space can be divided into three challenges: selecting a strategy
for log entry addressing, choosing a native I/O interface for storing log
entry content, and implementing efficient metadata management.

\begin{enumerate}
    \item {\bf Entry addressing.} We refer to the method by which a client
        locates a log entry in Ceph as entry addressing, and we consider two
        strategies. In a one-to-one (1:1) strategy each log entry is stored in
        a distinct object with a name derived from the associated log entry
        position. This is an attractive option because it is trivial for
        clients to locate a log entry given its position.  In contrast, an N:1
        strategy \emph{stripes} log entires across a smaller set of objects,
        but this adds complexity to both the client and the object interface which
        must multiplex a set of entries.

    \item {\bf Log entry storage.} Clients read and write binary data
        associated with each log entry, and these entries can be stored in the
        bytestream or in the key-value database associated with an object.
        Retrieval of log entry payloads should perform well for both large
        (e.g. database checkpoint) and small log entries.

    \item {\bf Metadata management.} The CORFU protocol defines the storage
        interface semantics, such as enforcing up-to-date epoch values and a
        write-once address space. The object interface constructed in Ceph must
        implement these semantics in software by storing metadata (e.g. the
        current epoch) and validating requests against this metadata (e.g. has
        the target position been written?). A key-value store is a natural
        location for this type of data, but metadata management adds overhead
        to each request and must be carefully designed.
\end{enumerate}

In the remainder of this section we will explore the full design space defined
by the cross product of these design challenges to arrive at a final design.

\subsection{Baseline Performance}

We begin the process of exploring the design space by exploiting the fact that
metadata management is a complexity and performance overhead that a full
implementation must incur beyond the costs of storing log entry data.
Therefore we first explore the design space by restricting the
space to log entry addressing and log entry storage (the \emph{I/O} action
shown in Figure~\ref{fig:corfu-sm}).  These two dimensions are
represented by the first two columns of Table~\ref{tab:pd-map} which describes
the entire design space. In the I/O column \emph{KV} corresponds to the
key-value interface, and the bytestream interface is represented by \emph{AP}
for an append strategy, and \emph{EX} for a strategy that writes to an
explicit bytestream offset (both described shortly).

\begin{table}[h]
\begin{tabular}{ | l | l | l | l | l |}
\hline
Map & I/O & Entry Size & Addressing & Metadata \\ \hline
\multirow{3}{*}{1:1} & KV  & Flex     & Ceph      & KV/BS \\ \cline{2-5}
                     & AP  & Flex     & Ceph/VFS  & KV/BS \\ \hline
\multirow{4}{*}{N:1} & KV  & Flex     & \multicolumn{2}{|c|}{KV/BS} \\ \cline{2-5}
                     & EX  & Fixed    & VFS       & KV/BS \\ \cline{2-5}
                     & AP  & Flex     & KV/BS     & KV/BS \\
\hline
\end{tabular}
\caption{The high-level design space of mapping CORFU log entry storage onto
the RADOS object storage system.}
\label{tab:pd-map}
\end{table}

Figure~\ref{fig:vanilla_wr_jewel} shows the expected performance of 1K log
appends without metadata management overhead using each of the five strategies
defined in Table~\ref{tab:pd-map}. The first thing to notice in this figure is
that both one-to-one strategies have relatively poor performance. The large
period of reduced throughput for \emph{11wr} corresponds to the OSD splitting
file system directories in order to maintain a maximum directory size, and
will occur in \emph{11kv} although the threshold number of objects is not
reached in this example due to reduced overall throughput of the \emph{11kv}
strategy.

The second lesson that we can learn from Figure~\ref{fig:vanilla_wr_jewel} is
that even when using an N:1 addressing strategy, the key-value interface
imposes a large overhead. This is unfortunate because the key-value interface
can provide a direct solution to addressing log entries within an object.
Instead, what we find is that an N:1 addressing strategy that stores log
entries in the bytestream using either object appends or writes to explicit
offsets outperform all other strategies by a factor of over 2x.

The apparent performance tie between the strategies of appending to an object
and writing to explicit offsets can be broken by considering the flexibility
offered by each approach. The third column \emph{entry size} in
Table~\ref{tab:pd-map} shows if a particular strategy supports storage of
entries with dynamic sizes, or if entries must be restricted to a fixed size.
Notably the strategy that stores log entries at explicit object offsets is
limited in this regard because each log entry is effectively pre-mapped into
the storage system. This leaves the clear winner: an N:1 strategy that appends
log entries to objects provides flexibility and the best write performance in
this particular configuration.  This result is corroborated by considering the
read performance for each strategy as well. Figure~\ref{fig:vanilla_rd_jewel}
shows the expected random read performance from a log containing 1K entries in
which an addressing strategy that stores log entries in the bytestream has the
best performance.

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{.47\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{experiments/librados-sweep/jewel_ssd.png}
      \caption{1hr,1k,wr,jewel}
      \label{fig:vanilla_wr_jewel}
  \end{subfigure}
  \begin{subfigure}[b]{.47\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-rand-read/output.read.60min.1k.png}
      \caption{1hr,1k,rd,jewel}
      \label{fig:vanilla_rd_jewel}
  \end{subfigure}
  \begin{subfigure}[b]{.47\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{experiments/basic-cls-overhead/output.1024.soft.reset.png}
      \caption{1hr,1k,cls\_wr,jewel}
      \label{fig:cls_wr_jewel}
  \end{subfigure}
  \begin{subfigure}[b]{.47\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{experiments/librados-sweep/firefly_ssd.png}
      \caption{1hr,1k,wr,firefly}
      \label{fig:vanilla_wr_firefly}
  \end{subfigure}
  \caption{placeholder.}
\end{figure}

\subsection{Metadata Management}

The previous results show that storing log entries in the bytestream has the
potential to provide the best overall performance, but by design, those
results do not contain the real-world overheads introduced by metadata
management such as validating that requests are tagged with an up-to-date
epoch value. Having only focused on entry I/O costs, in this section we
consider the overhead of the remaining actions shown in
Figure~\ref{fig:corfu-sm}.

\begin{enumerate}
    \item {\bf Epoch guard.} Each client request must be validated against the
        current epoch. This singleton value in infrequently updated, but the
        cost of accessing the value must be incurred for every request.

    \item {\bf Per-entry metadata.} Unlike the singleton epoch value, metadata
        associated with each log position must be read, and optionally updated
        for every request. For instance the \emph{fill} operation must ensure
        that it is not applied to a log position that has already been
        written. When an operation that changes a log position is successful
        (i.e. \emph{write}, \emph{fill}, \emph{trim}) the per-entry metadata
        must also be updated to reflect the change. While the size of the
        per-entry metadata is small, it is accessed for every operation and a
        single object may manage a large number of entries.
\end{enumerate}

The design space for managing metadata can be large depending on the
application. In many cases this space is significantly reduced when the
key-value interface is used because it handles a large portion of common data
management challenges such as indexing. On the other hand, we have seen that
the bytestream can be used to achieve much higher performance when used in
lieu of the key-value database, but leaves the design space wide open.

We consider two high-level design prototypes in this paper based on both the
key-value and bytestream interfaces. The summary of performance of these
designs is shown in Figure~\ref{fig:cls_wr_jewel} where we begin by
highlighting two baseline throughputs labeled \emph{librados} and
\emph{cls\_null} that correspond to the baseline append throughput described
in the previous section, and the append throughput achieved with \emph{object
class} overhead included. We will discuss the performance of our two prototype
designs relative to the base cost of using \emph{object class} facility.

The first design is based on the key-value interface. In this design the epoch
value is stored under a fixed key, and the metadata associated with each log
entry is stored under a key derived from the log position. The expected throughput of
this design is shown in Figure~\ref{fig:cls_wr_jewel} and labeled as
\emph{cls\_full\_omap}. This result shows us that even for very small values
the cost of managing data in the key-value store introduces significant
overhead. In contrast, the unstructured bytestream interface imposes little
to no restriction on how it is used.

An obvious choice for encoding the singleton epoch value in the bytestream is
to position it at a fixed offset such as in a header. A method for indexing
entry metadata is far less obvious. Solutions span a wide design spectrum and
include approaches such as encoding a search tree into the bytestream, or
exploiting the regularity log addressing and storing a dense array that can be
efficiently indexed. While there is clearly a wide variety of approaches, we
have chosen to examine the overhead incurred by an hypothetical best-case
design in which per-entry metadata can be accessed with a single I/O. The I/O
overhead of the test includes (1) read epoch value from a fixed header
location, (2) read the per-entry metadata, and in the case of an update (3)
write the updated per-entry metadata. The expected throughput from such a
design is shown in Figure~\ref{fig:cls_wr_jewel} labeled as
\emph{cls\_full\_hdr} and performs with roughly a 6\% overhead.

Application-level designs also contribute to an expanded design space.  In the
N:1 design of mapping CORFU onto Ceph the potential amount of I/O parallelism
is controlled by the value N, however if this value becomes too large there may
be no benefit or performance could suffer from reduced data locality. Other
aspects of application-level design are even more flexible. A wide variety of
data structures could be serialized into the byte stream to handle metadata
management tasks, but selecting a single external index that performs well
across workloads and hardware may not be possible.

\subsection{Cluster-Specific Optimizations}

In this section we have demonstrated that a design based on an N:1 addressing
scheme that stores both log entries and metadata in the bytestream can provide
the best overall performance, and it does so by a large margin. However, this
process of design we have outlined may not yield such clear cut results when
applied in a different context. To illustrate the complexity and breadth of
this design space, we describe the software, system tunables, and hardware
parameters that affect the performance of ZLog.

\subsubsection{Software Parameters}

Ceph releases stable versions every year (Oct/Nov) and LTS versions every 3-4
months~\cite{website:ceph-releases}. The head of the master branch moves
quickly because there are over 400 contributors and an active mailing list.
Over the calendar year, there were between  70 and 260 commits per
week~\cite{website:ceph-github}. These releases contain performance fixes and
new features that could directly ZLog.

\textbf{Ceph Versions}: to show the affect that this has on ZLog's design, we
ran the same benchmark that measures expected append throughput shown in
Figure~\ref{fig:vanilla_wr_jewel} on identical hardware but instead use a
long-term support version of Ceph released two years ago. The result of this
benchmark is shown in Figure~\ref{fig:vanilla_wr_firefly} in which the
bytestream interface outperforms the key-value interface by 12\%, compared to
over 100\% in the newer version of Ceph. Qualitatively, given the reduced
overall throughput achieved in the older version, some developers may find that
incurring the overhead of using the key-value interface is an easy decision
given the reduced complexity of the design space for metadata management. If
this choice had been made, then a future system upgrade could drop a
significant number of iops on the floor.

\textbf{Ceph Features}: to show the affect of cutting-edge Ceph features, we
now show results for BlueStore~\cite{weil:vault2016-bluestore}. BlueStore is a
replacement for the FileStore file system in the OSD (traditionally XFS).
FileStore has performance problems with transactions and enumerations; namely
the journal needed to assure atomicity incurs double writes and the file system
metadata model makes object listings slow, respectively. BlueStore stores data
directly on a block device and the metadata in RocksDB, which is provided by a
minimalistic, non-POSIX C++ filesystem. This model adheres to the overall
software defined storage strategy of Ceph because it gives the administrator
the flexiblity to store the 3 componenets of BlueStore (e.g., data, RocksDB
database, and RocksDB write-ahead log) on any partition on any device in the
OSD.

\subsubsection{System Tunables}

The most recent version of Ceph (v10.2.0-1281-g1f03205) has 994 tunable
parameters\footnote{This number comes from \texttt{src/common/config\_opts.h}
with debug options filtered out.}, where 195 of them pertain to the OSD itself
and 95 of them focus on the OSD back end file system (i.e. its
\texttt{filestore}). Ceph also has tunables for the subystems it uses, like
LevelDB (10 tunables), RocksDB (5 tunables), its own key-value stores (5
tunables), its object cache (6 tunables), its journals (24 tunables), and its
other optional filestores like BlueStore (49 tunables).

This many domain-specific tunables makes it almost impossible to come up with
the best set of tunables, although auto-tuning like the work done
in~\cite{behzad:sc2013-autotuning} could go a long way. Regardless of the
technique that we use, it is clear the the number of tunables increases the
physical design parameters to an unwieldy stae space size.

\subsubsection{Hardware Parameters}

Ceph is designed to run on a wide variety of commodity hardware as well as new
NVMe devices, all of which their own set of characteristics and tunables (e.g.,
the IO operation scheduler type). In our experiments, we tested SSD, HDDs, NVMe
devices, and

\subsection{Discussion}

This physical design process, of first thinking about the design trade-offs and
then doing parameters sweeps, is insufficient when general-purpose systems. Had
we built ZLog from the ground up (like CORFU) the story might be different but
to deal wtih the large state space we need a way to automatically and
autonomically decide which parameters to choose.

Next we present a declarative programming model that seeks to solve the issue
of selecting an optimal mapping between application requirements and object
class implementation.

\section{Programming Model}

Ceph is turning into a target for building distrubted services, bringing
distributed programming into the hands of many new people that may have
worked on the periphery of ceph.

Getting consistency and coordination correct is a challenge And as we have seen
optimization at a low-level given a simple set of semantics can yield a large
state space from which to draw designs Difficulty and challenges of writing
distributed software.  Programming distributed systems and applications.

We draw upon the recent work of Bloom a declarative lang thing

BLL (a.k.a. BRADOS) is a declarative language with relational (or tabular)
semantics for RADOS objects, which corresponds to a subset of the BLOOM
language (leaving out temporal logic). BLL borrows from BLOOM the table and
interface collections, the instantaneous implication (‘<=’) operator, as well
as all the collection operators.

BRADOS models the storage system state uniformly as a collection of relations.
The composition of a collection of existing interfaces is then expressed as a
collection of high-level \emph{queries} that describe how a stream of requests
(API calls) are filtered, transformed and combined with existing state to
define streams of outputs (API call returns as well as updates to system
state).  Separating the relational semantics of such compositions from details
of their physical implementations introduces a number of degrees of freedom,
similar to the ``data independence’’ offered by database query languages.  The
choice of access methods (for example, whether to use a bytestream interface or
a key/value store), storage decide classes(e.g., whether to use HDDs or SDDs),
physical layout details (e.g. a 1:1 or 1:N mapping) and execution plans (e.g.
operator ordering) can be postponed to execution time.  The optimal choices for
these physical design details are likely to change much more often than the
logical specification, freeing the interface designer from the need to rewrite
systems and device and interface characteristics change.

\subsection{The CORFU Interface in Bloom}

Listing~\ref{lab} shows code.

\begin{lstlisting}[caption={Hello}, label=lab]
class Corfu
include Bud

state do
table :epoch, [:epoch]
table :log, [:pos] => [:state, :data]

scratch :valid_op, [:type, :pos, :epoch] => [:data]
scratch :invalid_op, [:type, :pos, :epoch] => [:data]

interface input, :op, [:type, :pos, :epoch] => [:data]
interface output, :ret, [:type, :pos, :epoch] => [:retval]
end
\end{lstlisting}

\subsection{Optimizer}

\subsection{Other Interfaces}

The subset of the Bloom language we are using is powerful enough to express the
CORFU language, as well as many of the object class interfaces that are already
included in Ceph. We briefly sketch the Bloom imlementation of an object class
that provides a locking service. This is an interesting example because Bloom
implementations use mechanisms for introducing points of coordination in
execution, which in the context of Ceph, may be written in Bloom itself.

\begin{lstlisting}[caption={Hello}, label=lab]
bloom :take_lock do
end

bloom :break_lock do
end

bloom :list_locks do
end

bloom :lock_info do
end
\end{lstlisting}

\section{Evaluation}

\section{Related Work}

{\bf Active storage.}
There is a wide variety of work related \emph{active storage} that seeks to
increase performance and reduce data movement by exploiting remote storage
resources. This concept has been applied in the context of on-disk
controllers and object-based storage
(T10)~\cite{riedel:vldb98,du:nwesp05,xie:msst11}.  Solutions to safety
concerns have examined using managed lanuages, sandbox technologies, as well
as restricting extension installation to
vendors~\cite{john:hiperio08,xie:msst11,runde:msst12}.

Techniques for remote storage processing have been applied in
application-specific domains for database and file system acceleration, as
well as in remote data filtering in cloud-based storage
environments~\cite{uysal:hpca00,chiu:iccs03,lim:msst08,gkantsidis:nsdi13}
Others have collected and used statistics to optimize the location of
computation in general workloads as well as database systems
~\cite{chen:cluster12,chen:icpp12,qiao:icde08}.

Most closely related to our work are efforts that consider specific
programming models in the context of active storage such as a stream-based
model~\cite{acharya:asplos98}, a model that assisted in optmiizing balancing
computation~\cite{wickremesinghe:hpdc022}, and optimization that took
advantage of read-only data to reorder operations~\cite{huston:fast04}.
While our goals are similar to previous work, we start with a declarative
language that will enable us to apply optimization techniques beyond
the limited domain-specific optimizations found in previous work.

\section{Conclusion and Future Work}

Include performance management topic in future work.

\bibliography{paper}
\bibliographystyle{plain}

\end{document}